Spin canonical rings of log stacky curves

### 2015-07-07 curves-write-up/spin_cring.tex
minor reword before statement of main thm
rewording first 2 pgraphs of new ex & typographical

These bounds are far from tight and do not collectively
cover all cases in all genera. The main theorem of this paper
gives significantly tighter bounds for any log spin canonical ring on a log spin curve.

@@

\begin{example}
\label{eg:adding-point}
In this example, we shall compute the log canonical ring of $(\sx, 3 \infty, L),$ where $\sx$ has coarse space $X$, $\sx$ has signature $(0;3;3),$ and $\infty$ is a fixed closed point of $X,$ different from the stacky point $P_1$ with stabilizer order 3.

Following the technique which will later be introduced in Lemma ~\ref{lem:sat-1}, we deduce the structure of $R_L$ from the structure of the canonical ring of $R_{L_X} := R(X,3 \infty, L_X).$

Note that $R_{L_X} \cong k[x_1,x_2]$ where $\deg x_1 = \deg x_2 = 1.$ To see this, observe that by Riemann--Roch $h^0(X, L_X) = 2,$ so we will need two generators in degree $1$. Because $L_X$ is very ample, $R_{L_X}$ is generated in degree 1. To conclude, note that, $R_{L_X}$ does not have any relations. If there were some relation, then $\dim \proj R_{L_X} < 1$, contradictig the fact that $L_X$ is very ample, so $\proj R_{L_X} \cong X,$ which has dimension 1.

Next, we construct generators and relations for $R_L,$ using those of $R_{L_X}$. Note that we have a natural inclusion $\iota:R_{L_X} \hookrightarrow R_L$ induced by the inclusions $H^0(\sx,L_X) \hookrightarrow H^0(\sx,L).$ By Riemann Roch, we see there is some element $y_{1,3} \in (R_L)_3$ with $\ord_{P_1}(y_{1,3}) = -1,$ not in the image of the inclusion $\iota$. We claim that there exist $a_1, a_2 \in k$ and a degree 4 polynomial $f(x_1, x_2) \in k[x_1,x_2]$ such that
\begin{align*}
	R_L \cong k[x_1,x_2,y_{1,3}]/(a_1 x_1 y_{1,3} +a_2 x_2 y_{1,3} + f(x_1,x_2))
\end{align*}
First, note that $x_1,x_2,y_{1,3}$ generate all of $R_L$ from the Generalized Max Noether Theorem from Voight
and Zureick-Brown \cite[Lemma 3.1.1]{vzb:stacky}. That is, the maps
\begin{align*}
	H^0(\sx,3L) \otimes H^0(\sx,(n-3)L) \rightarrow H^0(\sx, nL)
\end{align*}
are surjective for $n \geq 4$. A relation of the form $a_1 x_1 y_{1,3} +a_2 x_2 y_{1,3} + f(x_1,x_2)=0$ must exist because $h^0(\sx,4L ) - h^0(\sx,4L_X) = 1,$ but $x_1 y_{1,3}$ and $x_2 y_{1,3}$ define two 
linearly independent elements with nontrivial image in the 1-dimensional vector space $H^0(\sx,4L )/ H^0(\sx,4L_X).$ So, we obtain a surjection
\begin{align}
\label{align:surjection}
k[x_1,x_2,y_{1,3}]/(a_1 x_1 y_{1,3} +a_2 x_2 y_{1,3} + f(x_1,x_2)) \rightarrow R_L.
\end{align}
To complete the example, it suffices to show there are no additional relations. First, note that $a_1 x_1 y_{1,3} +a_2 x_2 y_{1,3} + f(x_1,x_2)$ is irreducible because there are no relations among $x_1, x_2$ and $y_{1,3}$ in lower degrees.
Hence,  $k[x_1,x_2,y_{1,3}]/(a_1 x_1 y_{1,3} +a_2 x_2 y_{1,3} + f(x_1,x_2))$ is integral and is 2 dimensional. Thus, the map \eqref{align:surjection} defines a surjection from an integral 2 dimensional ring to a 2 dimensional ring. Therefore, it is an isomorphism.

\end{example}

>>>>

These bounds are far from tight and do not collectively
cover all cases in all genera. The main theorem of this paper
gives significantly tighter bounds for the log spin canonical ring
of any log spin curve.

@@

\begin{example}
\label{eg:adding-point}
Let $\sx$ be a stacky curve with coarse space $X$ and signature $(0
; 3; 3)$. Let $P_1$ denote the lone stacky point which has stabilizer
order $3$ and suppose $\infty$ is a fixed closed point of $X$ that is
not equal to $P_1$.

We will deduce the structure of the log spin canonical ring $R_L$
from the structure of the spin canonical ring $R_{L_X} := R(X, 3
\cdot \infty, L_X)$. This technique will later be generalized
and encapsulated in the statement and proof of Lemma ~\ref{lem:sat-1}.

Note that $R_{L_X} \cong k[x_1, x_2]$ where $\deg x_1 = \deg x_2 = 1.
$ To see this, observe that we will need two generators in degree $1$
because $h^0(X, L_X) = 2$ by Riemann--Roch. Because $L_X$ is very 
ample, we have that $R_{L_X}$ is generated in degree 1. To
conclude, note that $R_{L_X}$ does not have any relations. If there
exists some relation, then $\dim \proj R_{L_X} < 1$. This would
contradict the fact that $L_X$ is very ample. Thus, $\proj R_{L_X}
\cong X$ which has dimension 1.

Next, we construct generators and relations for $R_L$ using those
of $R_{L_X}$. Note that we have a natural inclusion $\iota: R_{L_X}
\hookrightarrow R_L$ induced by the inclusions $H^0(\sx, L_X)
\hookrightarrow H^0(\sx, L).$ By Riemann Roch, we see there is some
element $y_{1, 3} \in (R_L)_3$ with $\ord_{P_1}(y_{1, 3}) = -1,$ not
in the image of the inclusion $\iota$. We claim that there exist $a_
1, a_2 \in k$ and a degree 4 polynomial $f(x_1, x_2) \in k[x_1, x_2]$
such that
\begin{align*}
	R_L \cong k[x_1,x_2,y_{1,3}]/(a_1 x_1 y_{1,3} +a_2 x_2 y_{1,3} + f(x_1,x_2))
\end{align*}

First, note that $x_1, x_2, y_{1,3}$ generate all of $R_L$ from the
Generalized Max Noether Theorem from Voight and Zureick-Brown
\cite[Lemma 3.1.3]{vzb:stacky}. That is, the maps
\begin{align*}
	H^0(\sx, 3L) \otimes H^0(\sx, (n - 3)L) \rightarrow H^0(\sx, nL)
\end{align*}

\noindent
are surjective for $n \geq 4$. A relation of the form $a_1 x_1 y_{1,
3} +a_2 x_2 y_{1, 3} + f(x_1, x_2) = 0$ must exist because $h^0(\sx
, 4L) - h^0(\sx, 4L_X) = 1,$ but $x_1 y_{1, 3}$ and $x_2 y_{1, 3}$
define two linearly independent elements with nontrivial image in
the 1-dimensional vector space $H^0(\sx, 4L) / H^0(\sx, 4L_X).$ So,
we obtain a surjection
\begin{align}
\label{align:surjection}
	k[x_1,x_2,y_{1,3}]/(a_1 x_1 y_{1,3} +a_2 x_2 y_{1,3} + f(x_1,x_2)) 
	\rightarrow R_L.
\end{align}

To complete the example, it suffices to show there are no
additional relations. First, note that $a_1 x_1 y_{1, 3} + a_2 x_2
y_{ 1, 3} + f(x_1, x_2)$ is irreducible because there are no
relations among $x_1, x_2$ and $y_{1, 3}$ in lower degrees. Hence,
$k[x_1, x_2, y_{1, 3}] / (a_1 x_1 y_{1, 3} + a_2 x_2 y_{1, 3} +
f(x_1, x_2))$ is integral and is 2-dimensional. Thus, the map
\eqref{align:surjection} defines a surjection from an integral
2-dimensional ring to a 2-dimensional ring. Therefore, it is an 
isomorphism.
\end{example}
